---
title: "Reproducible Research: Peer Assessment 1"
author: "Ilana Golbin"
output: 
  html_document:
    keep_md: true
---
#Activity Monitoring Data
Ilana Golbin

## Loading and preprocessing the data
```{r loadData, echo=TRUE}
data<- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r histogram, echo=FALSE}
tapply_data <- tapply(data$steps, data$date, sum)
barplot(tapply_data)
title("Total Steps per Day")
# calculate mean and median
mean <- mean(tapply_data, na.rm=TRUE)
median <- median(tapply_data, na.rm=TRUE)
```

The median steps is `r median` and mean steps is `r mean`. 

## What is the average daily activity pattern?

```{r timeseries, echo=FALSE}
meanInterval_data <- tapply(data$steps, data$interval, mean, na.rm=TRUE)
plot(y=meanInterval_data, x=names(meanInterval_data), ylab="Steps", xlab="Interval")
title("Steps per Interval")
# calculate maximum
result <- which.max(meanInterval_data)
interval <- names(result)
```

The maximum steps occur in interval `r interval`.


## Imputing missing values
```{r naCount, echo=FALSE}
countNA <- sum(is.na(data))
# fill in missing data with mean for the interval

newData <- data

for (i in 1:length(newData$steps)) {
  
  if (is.na(newData$steps[i])) {
    missingInterval <- toString(newData$interval[i])
    meanValue <- as.numeric(getElement(meanInterval_data, missingInterval))
    newData$steps[i] <- meanValue
  }
  
}

# histogram of new dataframe
tapply_dataNew <- tapply(newData$steps, newData$date, sum)
barplot(tapply_dataNew)
title("Total Steps per Day, Adjusted")
# updated mean and median calculations
meanUpdated <- mean(tapply_dataNew)
medianUpdated <- median(tapply_dataNew)
```

The total number of missing values is: `r countNA`. The updated mean is `r meanUpdated` and updated median is `r medianUpdated`.

## Are there differences in activity patterns between weekdays and weekends?


```{r patterns, echo=FALSE}
#dayOfWeek <- factor("weekday", "weekend")
date <- data$date
date <- as.Date(date)
weekVec <- vector()
dayOfWeek <- vector()

for (d in 1:length(date)) {
  day <- weekdays(date[d])
  weekVec <- c(weekVec, day)
  if (day == "Saturday" | day == "Sunday") {
    dayOfWeek <- c(dayOfWeek, "weekend")
  } else {
    dayOfWeek <- c(dayOfWeek, "weekday")
  }
}
fact <- factor(dayOfWeek, labels=c("weekend", "weekday"))
factDATA <- cbind(data, fact) 

library(lattice)
attach(factDATA)
xyplot(steps~interval|fact)

```